Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-26 paper/paper.tex
minor changes after revising section 5.1.1.

\begin{table*}[t!]
  \small
  \caption{\label{table:attacks}List of attacks.\Mar{adicionar coluna ``Source''}}
  \centering
  \begin{tabular}{llllp{11cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Source} &
    \multicolumn{1}{c}{Category} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth (relevant/checked options)} \\
    \midrule     
    Ping Scan & NMAP & Active Recon. & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
    SYN Flood & hping3 & DoS & TCP & \textbf{flags}: S,12; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\

    JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
    %% WeBid \MyComment{Local File Inclusion} & RFI & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
    ColdFusion \MyComment{Privilege Escalation} & Nikto2 & Privilege Escalation & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
    .htaccess access\MyComment{attempted-recon} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
    .idq access\MyComment{EXPLOIT ISAPI} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
    iisadmin access\MyComment{web-application-attack} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

\subsubsection{Attacks.}Table~\ref{table:attacks} shows the attacks we considered in our
evaluation. These attacks cover different protocols and features of
\tname{}. Column ``Name'' shows the name of the attack, column
``Protocol'' shows the protocol name, and column ``Ground Truth''
shows the options of the rule that captures the attack.\Mar{Explicar o
  rationale para esta selecao de ataques!}

@@

\subsubsection{Malicious traffic.}~For some of the attacks listed on
Table~\ref{table:attacks}, we obtained pcap files directly from
trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}). Recall that \tname\ takes pcap files on
input. As explained on Section~\ref{sec:minimization}, these files
abstract network messages~\cite{pcap} and can be used to avoid network
traffic during rule synthesis. In addition to these sources, we used
the Nikto2~\cite{nikto} open source security tool as follows to obtain
pcap files for http and tcp attacks. We configured an http server
(Apache2) and ran Nikto2 to attack that server. Simultenously, we ran
WireShark~\cite{wireshark-net-monitor} to record the network traffic
and Suricata to analyze the traffic. We configured \suri\ with all
rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
default parameters, which makes it replay all recorded attacks. After
execution, we inspected the \suri\ log to identify which rules have
been activated and which packets were involved on each
activation. With the packet numbers, it is possible to recover the
corresponding pcap files, using Wireshark, as to ran \tname{} on
them. It is worth noting that some rules are triggered multiple times,
indicating that there are multiple instances of the same attack
encoded in Nikto2\Mar{@Lucas, diga exatamente como vc. obtem as
  variacoes. isto eh que eh importante para a proxima sentenca.}. RQ3
evaluates the importance of these variants to \tname{}'s
performance. Finally, we used the hping3~\cite{hping3} to reproduce
the Ping Scan icmp attack (see Section~\ref{sec:dos}) as Nikto2 only
supports tcp and http attacks. More precisely, we used the following
hping3 command to reproduce that attack: \CodeIn{hping3 -d 80 -w 512
  -S -p 80 --flood --rand-source <IP>}.

@@

>>>>

\begin{table*}[t!]
  \small
  \caption{\label{table:attacks}List of attacks.}
  \centering
  \begin{tabular}{llllp{10cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Source} &
    \multicolumn{1}{c}{Type} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth (relevant/checked options)} \\
    \midrule     
    Ping Scan & NMAP & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
    SYN Flood & hping3 & DoS & TCP & \textbf{flags}: S,12; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\

    JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
    %% WeBid \MyComment{Local File Inclusion} & RFI & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
    ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
    .htaccess access\MyComment{attempted-recon} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
    .idq access\MyComment{EXPLOIT ISAPI} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
    iisadmin access\MyComment{web-application-attack} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

\subsubsection{Attacks.}~Table~\ref{table:attacks} shows the selection
of attacks considered in our evaluation.  Column ``Name'' shows the
name of the attack, column ``Source'' shows the source where we
obtained data to reproduce the attack (see
Section~\ref{subsec:malicious-traffic}), column ``Type'' shows the
kind of attack, column ``Protocol'' shows the name of the protocol
explored in the attack, and column ``Ground Truth'' shows the rule
options to capture the attack. The rationale for the selection were
(1) to cover attacks with different characteristics, (2) to cover
attacks of various protocols, and (3) to explore different features of
\tname. As for (1), the selection includes attacks of the following
types: Denial-of-Service (DoS), Active Reconnaissance (AREC),
Privilege escalation (PESC), and Remote Code Execution (RCE). Note
that the distribution of attacks is not uniform across types and
protocols as they are not uniform in practice. For example, as
per~\cite{emerging-threats-open}, RCE are the most prevalent type of
attacks today.\Mar{please, confirm} Columns ``Protocol'' and ``Ground
Truth'' indicate, respectively, that we analyzed attacks of different
protocols and analyzed different features of \tname.

@@

\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~For some of the attacks listed on
Table~\ref{table:attacks}, we obtained pcap files directly from
trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}). Recall that \tname\ takes pcap files on
input. As explained on Section~\ref{sec:minimization}, these files
abstract network messages~\cite{pcap} and can be used to avoid network
traffic during rule synthesis. In addition to these sources, we used
the Nikto2~\cite{nikto} open source security tool as follows to obtain
pcap files for http and tcp attacks. We configured an http server
(Apache2) and ran Nikto2 to attack that server. Simultenously, we ran
WireShark~\cite{wireshark-net-monitor} to record the network traffic
and Suricata to analyze the traffic. We configured \suri\ with all
rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
default parameters, which makes it replay all recorded attacks. After
execution, we inspected the \suri\ log to identify which rules have
been activated and which packets were involved on each
activation. With the packet numbers, it is possible to recover the
corresponding pcap files, using Wireshark, as to ran \tname{} on
them. It is worth noting that some rules are triggered multiple times,
indicating that there are multiple instances of the same attack
encoded in Nikto2\Mar{@Lucas, diga exatamente como vc. obtem as
  variacoes. isto eh que eh importante para a proxima sentenca.}. RQ3
evaluates the importance of these variants to \tname{}'s
performance. We used NMAP~\cite{netmap} to reproduce the Ping Scan
attack (see Section~\ref{sec:active-recon}) and hping3~\cite{hping3}
to reproduce the SYN Flood attack (see
Section~\ref{sec:dos}).

@@

\Mar{move from above --> We used the following hping3 command to
  reproduce that attack: \CodeIn{hping3 -d 80 -w 512 -S -p 80 --flood
    --rand-source <IP>}.}
